---
title: "p8105_hw3_ag2965"
author: "Aakriti Shukla"
date: "2022-10-06"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Let's load the necessary libraries.
```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
data("instacart")
```


The goal is to do some exploration of the instacart dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?

There are `r n_distinct(pull(instacart,aisle_id))` aisles. The most items are ordered from XXX aisle.

```{r}
instacart%>%
  janitor::clean_names()%>%
  summarize(
    n_distinct(aisle_id))
```

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart%>%
  ggplot(aes(x=aisle_id))+
  geom_histogram()
```
filter(aisle_id>10000)

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart%>%
  ggplot(aes(x=aisle))+
  geom_histogram()
  filter(aisle_id>10000)
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

Problem 2
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

minutes (data)--> numeric
days--> factor
#Use factor whenever there are a limited number of variables to analyze (e.g. seasons, days of the week, most common names, etc)


    
```{r}
accel_data = read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    minutes=as.numeric(minutes),
    day_type= recode(day,"Monday"="weekday", "Tuesday"="weekday",  "Wednesday"="weekday", "Thursday"="weekday", "Friday"="weekday", "Saturday"="weekend", "Sunday"="weekend"))
```

##Describe resulting dataset ##Need to do this

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_data=read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    minutes=as.numeric(minutes),
    day_type= recode(day,"Monday"="weekday", "Tuesday"="weekday",  "Wednesday"="weekday", "Thursday"="weekday", "Friday"="weekday", "Saturday"="weekend", "Sunday"="weekend"))%>%
  group_by(day)%>%
  summarize(sum(minutes))
accel_data%>%knitr::kable(digit=2)
```

##Add trends 

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_data=read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    activity=as.numeric(activity),
    minutes=as.numeric(minutes),
    day_type= recode(day,"Monday"="weekday", "Tuesday"="weekday",  "Wednesday"="weekday", "Thursday"="weekday", "Friday"="weekday", "Saturday"="weekend", "Sunday"="weekend"))

accel_data%>%
  ggplot(aes(x=minutes,y=activity,color=day))+
  geom_smooth(se = FALSE)

head(accel_data)
```

#Problem 3

We will first load the necessary dataset.
```{r}
library(p8105.datasets)
data("ny_noaa")
```
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

The National Oceanic and Atmospheric Association (NOAA) 
id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C)

There are `r nrow(ny_noaa)`rows and `r ncol(ny_noaa)` columns in the dataset.
`r range(pull(ny_noaa,tmax))`

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?


```{r}
ny_noaa_mini=sample_n(ny_noaa, 1000)%>%
  janitor::clean_names()%>%
  separate(col=date,into=c('year','month','date'), sep="-")%>%
    mutate(
    month=as.integer(month),
    year=as.integer(year),
    month=replace(month, month == "1","January"),
    month=replace(month, month == "2","February"),
    month=replace(month, month == "3","March"),
    month=replace(month, month == "4","April"),
    month=replace(month, month == "5","May"),
    month=replace(month, month == "6","June"),
    month=replace(month, month == "7","July"),
    month=replace(month, month == "8","August"),
    month=replace(month, month == "9","September"),
    month=replace(month, month == "10","October"),
    month=replace(month, month == "11","November"),
    month=replace(month, month == "12","December"),
    prcp=as.numeric(prcp),
    snow=as.numeric(snow),
    snwd=as.numeric(snwd),
    tmax=as.numeric(tmax),
    tmin=as.numeric(tmin),
    prcp=prcp/100,
    snow=snow/10,
    snwd=snwd/10,
    tmax=tmax/10,
    tmin=tmin/10,
    )%>%
  arrange(id,year,month,date,prcp,snow,snwd,tmax,tmin)
```


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
ny_noaa%>%
  ggplot(aes(x=id,y=summarize(mean(tmax)),color=month))+
  geom_point()+
  filter(month=="January", "July")
```




Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

