---
title: "p8105_hw3_ag2965"
author: "Aakriti Shukla"
date: "2022-10-06"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Let's load the necessary libraries.
```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
data("instacart")
```

#Problem 1

The goal is to do some exploration of the instacart dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

How many aisles are there, and which aisles are the most items ordered from?

There are `r n_distinct(pull(instacart,aisle_id))` aisles. The most items are ordered from the fresh fruits and fresh vegetables aisles.

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)

instacart%>%
  janitor::clean_names()%>%
  summarize(
    n_distinct(aisle_id))
```

```{r}
instacart%>%
  count(aisle)%>%
  filter(n>100000)
```


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart%>%
  count(aisle)%>%
  filter(n>10000)%>%
  mutate(aisle=fct_reorder(aisle,n))%>%
  ggplot(aes(x=aisle, y=n))+
  geom_point()+
  labs(title = "Number of items ordered in each aisle",
       y="number of items") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart%>%
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits"))%>%
  group_by(aisle)%>%
  count(product_name)%>%
  mutate(rank=min_rank(desc(n)))%>%
  filter(rank<4)%>%
  arrange(desc(n))%>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacartapplescoffee=instacart%>%
   filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream"))%>%
  group_by(product_name,order_dow)%>%
  summarize(mean(order_hour_of_day))%>%
  mutate(
      order_dow=recode(order_dow,"0"="Sunday", "1"="Monday", "2"="Tuesday","3"="Wednesday","4"="Thursday","5"="Friday", "6"="Saturday"))%>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean(order_hour_of_day)"
  )%>%
 knitr::kable()

head(instacartapplescoffee)
```


#Problem 2
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.



```{r, show_col_types = FALSE}
accel_data = read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    activity=as.numeric(activity),
    minutes=as.numeric(minutes),
    day_type= recode(day,"Monday"="weekday", "Tuesday"="weekday",  "Wednesday"="weekday", "Thursday"="weekday", "Friday"="weekday", "Saturday"="weekend", "Sunday"="weekend"))

head(accel_data)
```


Above, I have loaded and tidied the data. The dataset includes all originally observed variables and values, has useful variable names, and includes a weekday vs weekend variable. The resulting dataset has `r nrow(accel_data)` rows and `r ncol(accel_data)` columns in the dataset. The variables include week number, day of the week, minute of the day, activity at each minute of the day, and whether the day of the week is a weekday or weekend. There are a total of 5 weeks in which the accelerometer is worn. The range of the activity variable is `r range(pull(accel_data,activity))`.

Below, I have wrangled the data to focus on the total activity over the day. Using my tidied dataset, I have aggregated across minutes to create a total activity variable for each day, and create a table showing these totals. This table suggests that activity levels tend to be higher on Fridays than on the other days of the week.

```{r}
accel_data_agg=read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    activity=as.numeric(activity),
    minutes=as.numeric(minutes),
    day_id=as.factor(day_id))%>%
    group_by(week,day)%>%
  summarize(sum(activity))%>%
  ungroup(week,day)%>%
   mutate(
        day = forcats::fct_relevel(day, "Sunday", "Monday", "Tuesday","Wednesday","Thursday","Friday","Saturday"))%>%
  pivot_wider(
    names_from="day",
    values_from="sum(activity)")%>%
knitr::kable(digit=2)

head(accel_data_agg)
```
 

Accelerometer data allows the inspection activity over the course of the day. Below, I have made a single-panel plot that shows the 24-hour activity time courses for each day. The plot suggests that activity levels are lowest overnight and have a bimodal distribution during the waking part of the day, with highest levels in the morning and evening and a slight dip in activity in the middle of the day.

```{r}
accel_data_24=read_csv("data/accel_data.csv")%>%
  janitor::clean_names()%>%
   pivot_longer(
    activity_1:activity_1440,
    names_to="minutes",
    values_to="activity",
    names_prefix="activity_")%>%
  mutate(
    day=as.factor(day),
    activity=as.numeric(activity),
    minutes=as.numeric(minutes),
    day_type= recode(day,"Monday"="weekday", "Tuesday"="weekday",  "Wednesday"="weekday", "Thursday"="weekday", "Friday"="weekday", "Saturday"="weekend", "Sunday"="weekend"))

accel_data_24%>%
  ggplot(aes(x=minutes,y=activity,color=day))+
  geom_smooth(se = FALSE)+
labs(
  x="Minute of Day",
  y="Activity Level",
  title="Activity Level by Time of Day",
  caption="Data came from accel_data package"
)

head(accel_data_24)
```

#Problem 3

We will first load the necessary dataset.
```{r}
library(p8105.datasets)
data("ny_noaa")
```
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

This data from The National Oceanic and Atmospheric Association (NOAA) includes the weather station ID, date of observation, precipitation in tenths of mm, snowfall in mm, snow depth in mm, and maximum and minimum temperatures in tenths of degrees Celsius. There are `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns in the dataset. There are many missing values  in this dataset.

Below, I have done some data cleaning. I have created separate variables for year, month, and day. I have changed the units of the observations for temperature to degrees Celsius, precipitation to centimeters,  snowfall to centimeters, snow depth to centimenters; these are now given in reasonable units. Based on the density plot created below, the most commonly observed value for snowfall was 0 because the majority of places did not have snow, or if they did, they did not have the same levels of snow.

```{r}
ny_noaa_tidy=ny_noaa%>%
  janitor::clean_names()%>%
  separate(col=date,into=c('year','month','date'), sep="-")%>%
    mutate(
    month=as.integer(month),
    year=as.integer(year),
    snow=as.numeric(snow),
    month=replace(month, month == "1","January"),
    month=replace(month, month == "2","February"),
    month=replace(month, month == "3","March"),
    month=replace(month, month == "4","April"),
    month=replace(month, month == "5","May"),
    month=replace(month, month == "6","June"),
    month=replace(month, month == "7","July"),
    month=replace(month, month == "8","August"),
    month=replace(month, month == "9","September"),
    month=replace(month, month == "10","October"),
    month=replace(month, month == "11","November"),
    month=replace(month, month == "12","December"),
    prcp=as.numeric(prcp),
    snow=as.numeric(snow),
    snwd=as.numeric(snwd),
    tmax=as.numeric(tmax),
    tmin=as.numeric(tmin),
    prcp=prcp/100,
    snow=snow/10,
    snwd=snwd/10,
    tmax=tmax/10,
    tmin=tmin/10,
    )%>%
  arrange(id,year,month,date,prcp,snow,snwd,tmax,tmin)

head(ny_noaa)

ggplot(ny_noaa, aes(x = snow))+ 
  geom_density(alpha = .4)
```


Below, I have made a two-panel plot showing the average max temperature in January and in July in each station across years. This plot demonstrates that the average maximum temperature is approximately 25 degrees Celsius warmer in July than January. It appears that the year 1988 had an exceptionally cold month of July, with a maximal temperature that was about 8 degrees Celsius lower than other years. There is some variation in temperatures each year, with an overall regression to the mean.

```{r}
ny_noaa_tidy%>%
  arrange(id, year, month)%>%
  group_by(month,year,id)%>%
  summarize(
    tmax_mean=mean(tmax,na.rm=TRUE))%>%
    filter(
      month%in% c("January","July"))%>%
ggplot(aes(x=year,y=tmax_mean, group=id))+
  geom_line(alpha=0.5)+
    facet_grid(.~month)+
labs(
  x="Year",
  y="Average Tmax",
  title="Average Tmax for January and July Across All Centers",
  caption="Data came from ny_noaa package"
)

```




Below, I have made a two-panel plot showing (i) a scatterplot showing tmax vs tmin for the full dataset and (ii) a density plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.The plots demonstrate that snowfall was less than 40 centimeters in most years.

```{r}
tmaxtminplot=ny_noaa_tidy%>%
  drop_na(c(tmax,tmin))%>%
ggplot(aes(x=tmin,y=tmax))+
  geom_hex()+
  labs(
  x="Tmax in degrees C",
  y="Tmin in degrees C",
  title="Tmax vs. Tmin Across All Centers",
  caption="Data came from ny_noaa package"
)

tmaxtminplot
```

```{r}
library(ggridges)
snow_fall=ny_noaa_tidy%>%
  mutate(snow=snow*10)%>%
  filter(0<snow,snow<100)%>%
ggplot(aes(x=snow,y=factor(year)))+
  geom_density_ridges(scale = .85)+
  labs(
  x="Snowfall in mm",
  y="Year",
  title="Snowfall by Year",
  caption="Data came from ny_noaa package"
)

snow_fall
```

```{r}
library(patchwork)
tmaxtminplot + snow_fall
```